% Created 2014-10-13 Mon 10:39
\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{fixltx2e}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{marvosym}
\usepackage{wasysym}
\usepackage{amssymb}
\usepackage{hyperref}
\tolerance=1000
\author{Clarissa Littler}
\date{\today}
\title{Homework 1}
\hypersetup{
  pdfkeywords={},
  pdfsubject={},
  pdfcreator={Emacs 24.3.1 (Org mode 8.2.7b)}}
\begin{document}

\maketitle

\section{HW 1: Regular Languages (Due 10/21/2014T)}
\label{sec-1}
\subsection{Problem 1}
\label{sec-1-1}
Give state diagrams of DFAs for the following languages
\begin{itemize}
\item $\{ w | w \text{ contains the substring} ab \text{ and } ba  \}$ ($\Sigma$ = \{a,b\})
\item $\{ w | w \text{ contains an even number of 0s or exactly three 1s} \}$ (hint: this is the union of two languages)
\item $\{ w | w \text{ is a binary multiple of 5} \}$ Note: there's a trick to this one. As a hint there should be a total of \emph{five} states in your DFA and only one accept state. As another hint, think about reading a binary number from left to right and how you calculate the number as an iterative process. Clarification: \textit{[2014-10-13 Mon]} Depending on if you want to accept the empty string in the language it could actually be 6 states. It's 5 states if you let the empty string count as the binary number 0, but you could also consider the empty string to be NaN.
\end{itemize}
\subsection{Problem 2}
\label{sec-1-2}
Prove or disprove the following: let $D$ be a DFA with $k$-states. If the language $L(D)$ is \emph{finite} then there exists at least one string $s$ of length at-most $k-1$ such that $D$ does not accept  $s$. Hint: what do we know about a DFA if its language is finite?
\subsection{Problem 3}
\label{sec-1-3}
For any string $w=w_1 w_2 \ldots w_n$ then $w^R = w_n w_{n-1} \ldots w_1$ is the reverse of the string. For any language $A$, let $A^R = \{ w^R | w \in A \}$. Prove that if $A$ is regular then so is $A^R$. 
\subsection{Problem 4}
\label{sec-1-4}
Give NFAs for the following languages 
\begin{itemize}
\item $\{ w | w \text{ uses an arbitrary number of all but one of } \{a,b,c,d\} \}$ (Clarification: for the purposes of this language strings such as $ab$, $a$, $\epsilon$ are all \textbf{in} the language. Basically, any string that doesn't use all four letters of the alphabet is in the language.)
\item $\{ w | w \text{ ends with a zero} \}$ but you must use only two states
\end{itemize}
\subsection{Problem 5}
\label{sec-1-5}
(from 1.38 in Sipser) An all-NFA M is a 5-tuple (Q, $\Sigma$, $\delta$, q$_{\text{0}}$, F) that is like an NFA accept that the acceptance condition is that \emph{every} possible state the NFA can be in at the end of processing the string must be an accept state. Prove that all-NFAs are equivalent in power to DFAs. Hint: what does the acceptance condition mean in terms of $\exists$ and $\forall$ and if there are \emph{no} possible states you can be in?
\subsection{Problem 6}
\label{sec-1-6}
Prove \emph{using the pumping lemma} that the following languages aren't regular. Remember, your proof needs to be a proof by contradiction in the form we did in class. 
\begin{itemize}
\item \{ 0$^{\text{n}}$ 1$^{\text{m}}$ | m > n\}
\item \{ s | s has an even number of 0s and fewer 1s than pairs of 0s \}
\end{itemize}
% Emacs 24.3.1 (Org mode 8.2.7b)
\end{document}
