% Created 2014-07-31 Thu 10:40
\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{fixltx2e}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{marvosym}
\usepackage{wasysym}
\usepackage{amssymb}
\usepackage{hyperref}
\tolerance=1000
\author{clarissa}
\date{\today}
\title{HW2}
\hypersetup{
  pdfkeywords={},
  pdfsubject={},
  pdfcreator={Emacs 24.3.1 (Org mode 8.2.7b)}}
\begin{document}

\maketitle
\tableofcontents

\section{HW 2: Context-Free Languages}
\label{sec-1}
\subsection{Problem 1}
\label{sec-1-1}
\begin{itemize}
\item Using the pumping lemma, show that the language $\{ w | w \text{ is not a palindrome} \}$ is not regular. Hint: we know that the regular languages are closed under complement.
\item Give a CFG for the language $\{ w | w \text{ is not a palindrome} \}$ (let $\Sigma$ be  \{0,1\})
\item Convert the CFG from the previous part into a PDA that decides the same language
\end{itemize}
\subsection{Problem 2}
\label{sec-1-2}
The languages $\{0^s1^s2^t | s,t \ge 0 \}$ and $\{0^s1^t2^t | s,t \ge 0 \}$ are context-free. 
\begin{itemize}
\item Write CFGs for each of them
\item By taking the intersection of these two languages, show that the context-free languages are \emph{not} closed under intersection. Use the context-free pumping lemma
\item Use the result of the previous part and DeMorgan's Laws to show that the context-free languages are \emph{alos} not closed under complement.
\end{itemize}
\subsection{Problem 3}
\label{sec-1-3}
Write a context-free grammar for \{ w | w is a palindrome \} that is in Chomsky Normal Form. Hint: it might be easiest to make the simplest possible grammar and then \emph{transform} it into CNF.
\subsection{Problem 4}
\label{sec-1-4}
Prove that if a $G$ is a CFG in Chomsky Normal Form then for any non-empty string $w \in L(G)$ of length $n$ then exactly $2n - 1$ steps are required to derive $w$. You don't have to prove this with induction, just give an argument based upon the restricted structure of Chomsky Normal Form. As a side note, we'll come back to this result much later in the course.
% Emacs 24.3.1 (Org mode 8.2.7b)
\end{document}
